---
title: "Question 2"
output:
  pdf_document: default
  html_document: default
---

#### Comparing two sub-populations

Consider the TitanicSurvival Data Set. The interest lies in comparing the *age* of the two sub-populations $P_1$: female passengers and $P_2$: male passengers on titanic. 

```{r, echo=FALSE}
library(carData)
data(TitanicSurvival)
Titanic = na.omit(TitanicSurvival)
Titanic$survived1 = as.numeric(Titanic$survived == "yes")
```
```{r}
Titanic.Female = Titanic[Titanic$sex == "female",]
Titanic.Female = Titanic.Female$age
Titanic.Male = Titanic[Titanic$sex == "male",]
Titanic.Male = Titanic.Male$age
pop = list(pop1 = Titanic.Female, pop2 = Titanic.Male)
```

We will now consider 3 discrepancy measures: 
$$ D_1(P_1,P_2) = \frac {\bar{Y_1} - \bar{Y_2}} {\tilde \sigma \sqrt{1/n_1 + 1/n2}} $$
$$ D_2(P_1,P_2) = \frac {SD(P_1)} {SD(P_2)} - 1$$
$$ D_3(P_1,P_2) = \frac {IQR(P_1)} {IQR(P_2)} - 1$$
We will write the necessary functions - mixRandomly, D1Fn, D2Fn, D3Fn (hidden due to space constraint).

```{r, echo=FALSE}
mixRandomly <- function(pop) {
  pop1 <- pop$pop1
  n_pop1 <- length(pop1)

  pop2 <- pop$pop2
  n_pop2 <- length(pop2)

  mix <- c(pop1,pop2)
  select4pop1 <- sample(1:(n_pop1 + n_pop2),
                     n_pop1,
                     replace = FALSE)

  new_pop1 = mix[select4pop1]  
  new_pop2 = mix[-select4pop1]
  list(pop1=new_pop1, pop2=new_pop2)
}

# Itroducing D1 as a function
D1Fn <- function(pop) {
    ## First sub-population
    pop1 <- pop$pop1
    n1 <- length(pop1)
    m1 <- mean(pop1)
    v1 <- var(pop1)
    
    ## Second sub-population
    pop2 <- pop$pop2
    n2 <- length(pop2)
    m2 <- mean(pop2)
    v2 <- var(pop2)
    
    ## Pool the variances
    v <- ((n1 - 1) * v1 + (n2 - 1) * v2)/(n1 + n2 - 2)
    
    ## Determine the t-statistic
    temp <- (m1 - m2) / sqrt(v * ( (1/n1) + (1/n2) ) )
    
    ## Return the t-value
    temp
}

# Itroducing D2 as a function
D2Fn <- function(pop) {
    ## First sub-population
    pop1 <- pop$pop1
    n1 = length(pop1)
    SD1 <- sqrt(var(pop1)*(n1-1)/n1)

    ## Second sub-population
    pop2 <- pop$pop2
    n2 = length(pop2)
    SD2 <- sqrt(var(pop2)*(n2-1)/n2)
    
    ## Determine the test statistic
    temp <- SD1/SD2 - 1
    
    ## Return the t-value
    temp
}

# Itroducing D3 as a function
D3Fn <- function(pop) {
    ## First sub-population
    pop1 <- pop$pop1
    IQR1 <- IQR(pop1)

    ## Second sub-population
    pop2 <- pop$pop2
    IQR2 <- IQR(pop2)
    
    ## Determine the test statistic
    temp <- IQR1/IQR2 - 1
    
    ## Return the t-value
    temp
}
```

Let us now generate the historgrams of the three discrepency measures based on 5000 shuffles on the two subpopulations $P_1$: female passengers on Titanic $P_2$: male passengers on Titanic. We will also superimpose the observed discrepency measure on these histograms. 

```{r}
par(mfrow = c(1,3))

# Plot for D1
D1Vals <- sapply(1:5000, FUN = function(...){D1Fn(mixRandomly(pop))})

hist(D1Vals, breaks=40, probability = TRUE, 
     main = "Permuted populations", xlab="D1 statistic",
     col="lightgrey")
abline(v=D1Fn(pop), col = "red", lwd=2)

#Plot for D2
D2Vals <- sapply(1:5000, FUN = function(...){D2Fn(mixRandomly(pop))})
hist(D2Vals, breaks=40, probability = TRUE, 
     main = "Permuted populations", xlab="D2 statistic",
     col="lightgrey")
abline(v=D2Fn(pop), col = "red", lwd=2)

#Plot for D3
D3Vals <- sapply(1:5000, FUN = function(...){D3Fn(mixRandomly(pop))})
hist(D3Vals, breaks=40, probability = TRUE, 
     main = "Permuted populations", xlab="D3 statistic",
     col="lightgrey")
abline(v=D3Fn(pop), col = "red", lwd=2)
```

We will now use all the three discrepancy measures $D_1, D_2, D_3$ to perform a multiple test to compare the two subpopulations $P_1$ and $P_2$. We will use $M = M^* = 300$ for multiple  testing. 

```{r, echo=FALSE}
#pop is a list whose two members are two sub-populations

#discrepancies is a list whose K members are K discrepancy measures: D1,...,Dk
calculateSLmulti <- function(pop, discrepancies, B_outer = 1000, B_inner){
  if (missing(B_inner)) B_inner <- B_outer
  ## Local function to calculate the significance levels
  ## over the discrepancies and return their minimum
  
  getSLmin <- function(basePop, discrepanies, B) {
  observedVals <- sapply(discrepancies, 
                           FUN = function(discrepancy) {discrepancy(basePop)})
    

    K <- length(discrepancies)

    total <- Reduce(function(counts, i){
      #mixRandomly mixes the two populations randomly, 
      #so the new sub-populations are indistinguishable
      NewPop <- mixRandomly(basePop)
      
      ## calculate the discrepancy and counts
      Map(function(k) {
        Dk <- discrepancies[[k]](NewPop)
        if (Dk >= observedVals[k]) counts[k] <<- counts[k] +1 },
        1:K) 
      counts
    }, 
    1:B, init = numeric(length=K)) 
    
    SLs <- total/B
    min(SLs)
  }
  
  SLmin <- getSLmin(pop, discrepancies, B_inner)
  
  total <- Reduce(function(count, b){
    basePop <- mixRandomly(pop)
    if (getSLmin(basePop, discrepancies, B_inner) <= SLmin) count + 1 else count
  },   1:B_outer, init = 0)
  
  SLstar <- total/B_outer
  SLstar
}
```

```{r}
discrepancies <- list(D1Fn , D2Fn , D3Fn)
### The following takes a long time (about 20 minutes)
### for B_outer = B_inner = 1,000 say
### So for illustration much smaller values than would be sensible are
### used here
set.seed(341)
SLstar=calculateSLmulti(pop, discrepancies, B_outer = 300, B_inner=300)
SLstar
```

#### Conclusion 
$\bullet$ Since the p-value/significance level is large ($\approx$ 0.63), there is no evidence against the hypothesis that the Male and Female passengers were randomly drawn from the same population based on the discrepency measures $D_1, D_2, D_3$ of age of passengers
