---
title: "Question 3"
output:
  pdf_document: default
  html_document: default
---

#### Bootstrap-t Confidence Intervals

The t-distribution approximation of a sampling distribution works for certain attributes usually when $\tilde{a}(S)$ is approximately normal over all possible samples. If $a(P)$ is the median or a measure of skewness, the t-distribution cannot be a good approximation 

In the Bootstrap-t Confidence Interval approach we use bootstrap to approximate the sampling distribution of a pivotal quantity and then construct a confidence interval based on it. This approach is different from that of t-approximation since we also require an estimate of the standard error of an attribute. 


#### How to calculate confidence interval of the pivotal quantity using bootstrap estimate? 

To use the bootstrap to approximate the sampling distribution of $Z$, we first estimate the population $P$ with the estimate $P^* = S$ (the sample). Then we estimate the sample $S$ with the bootstrap sample $S^*$ and generate $S^*_1,...,S^*_B$. We then calculate
$$ Z^*_b = \frac {\tilde{a}(S_b) - a(S)} {\hat{SD}[\tilde{a}(S_b^*)]}$$
From above the bootstrap estimate of the sampling distribution is $\{ z^*_1,...,z^*_b\}$.
Using a p $\in$ (0,1) the bootstrap estimate we can find $Z^*_{lower}$ and $Z^*_{upper}$ such that 
$$ 1-p = Pr(Z^*_{lower} \leq Z^* \leq Z^*_{upper}) \approx Pr(Z^*_{lower} \leq Z \leq Z^*_{upper})$$

\

And a confidence interval of the pivotal quantity using the above bootstrap estimate is
$$ (a(S) - Z^*_{upper} \times {\hat{SD}[\tilde{a}(S)]}, a(S) - Z^*_{lower} \times {\hat{SD}[\tilde{a}(S)]})$$

#### Example 1
We will now look at the spotify data and compare the bootstrap estimate to the sampling distribution for the average energy of each song. 

```{r echo=FALSE}
spotify <- read.csv("./spotify.csv", header=TRUE)
popSpotify <- rownames(spotify)
M <- 10^4
n = 6
set.seed(341)
sdn <- function( y.pop ) {
N = length(y.pop)
sqrt( var(y.pop)*(N-1)/(N) ) }
se.avg <- function(y.sam) {
  sdn(y.sam)/sqrt(length(y.sam)) * sqrt((65-length(y.sam))/(65-1) )
}
popSize <- function(pop) {
  if (is.vector(pop))
  {if (is.logical(pop))
## then assume TRUE values identify units
  sum(pop) else length(pop)}
else nrow(pop)
}
getSample <- function(pop, size, replace=FALSE) {
N <- popSize(pop)
pop[sample(1:N, size, replace = replace)]
}

samples <- sapply(1:M, FUN =function(m) sample(popSpotify, n, replace = FALSE)) # S1,..,SM
avePop <- mean(spotify[, "energy"])
avesSamp <- apply(samples, MARGIN = 2,
FUN = function(s){mean(spotify[s,"energy"])}) # a(Si)
SEaveSamp <- apply(samples, MARGIN = 2,
FUN = function(s){se.avg(spotify[s,"energy"])}) # SE^

samSpotify <- sample(popSpotify, n, replace = FALSE) # one sample S
samStar <- sapply(1:M, FUN = function(m) sample(samSpotify, n, replace = TRUE)) # s1*,..., sb*
aveSam <- mean(spotify[samSpotify, "energy"]) # a(S)
avesBoot <- apply(samStar, MARGIN = 2, FUN = function(s) { # a(Sb*)
  mean(spotify[s, "energy"])
})
SEaveBoot <- apply(samStar, MARGIN = 2, FUN = function(s) { # standard error
  se.avg(spotify[s, "energy"])
})
```
```{r}
ZPop <- (avesSamp - mean(spotify[,"energy"]))/SEaveSamp # Zi
ZBoot <- (avesBoot - aveSam)/SEaveBoot
```

The plots below compares both the approaches and we can see that the histogram and the $t_5$ line (in blue) are quite close

```{r}
savePar <- par(mfrow = c(1, 3))
brk = seq(-60, 60, by = 0.25)
hist(ZPop, freq = FALSE, breaks = brk, col = adjustcolor("grey", 0.5), main = paste("Sampling Distribution n=6"), xlab = "Z", xlim = c(-5,5), ylim = c(0, 0.5))  # based on many Sample
lines(x = seq(-4, 4, 0.1), y = dt(x = seq(-4, 4, 0.1), df = n - 1), col = "blue",
lwd = 2) # t5 line
hist(ZBoot, freq = FALSE, breaks = brk, col = adjustcolor("grey", 0.5), main = paste("Sampling Distribution n=6"), xlab = "Z", xlim = c(-5, 5), ylim = c(0, 0.5)) # based on one S and many bootstrap S
plot(ecdf(ZBoot), xlim = c(-4, 4), col = "red", main = "Compare ECDF", xlab = "Z",
ylab = "F(z)")
lines(ecdf(ZPop), col = "blue") 
```

**The General Approach to calculating Bootstrap-t confidence interval:**

\

For a given sample $S$, attribute $a(S)$ and standard error $\hat{SD}[\tilde{a}(S)]$.

* We first calculate $a(S)$ and  $\hat{SD}[\tilde{a}(S)]$ based on the sample. We then generate $B$ bootstrap samples $S^*_1,...,S^*_B$ from $S$ *with replacement*. 
* For each of the $B$ bootstrap samples from above, we then calculate
$a(S^*_b)$ and $\hat{SD}[\tilde{a}(S_b^*)]$ such that $$Z_b^*= \frac {a(S_b^*) - a(S)} {\hat{SD}[\tilde{a}(S_b^*)]}$$
* From the values $z_1^*,...,z_B^*$ we find $c_{lower} = Q_z (p/2)$ and $c_{upper} = Q_z (1 - p/2)$ 
The pair $\{ c_{lower},c_{upper} \}$ is nothing but quantiles from $\{ z_1^*,...,z_B^* \}$ which are estimates of the sampling distribution. 
* We can now find a 100$(1-p)\%$ bootstrap-t confidence interval, given by:
$$ ( a(S) - c_{upper} \times \hat{SD}[\tilde{a}(S)], a(S) - c_{lower}\times \hat{SD}[\tilde{a}(S)])$$

#### Example 2
Let's calculate the bootstrap-t confidence interval using the standard error $\hat{SD}[\tilde{a}(S)]$

```{r}
samSpotifyEnergy = spotify[samSpotify, "energy"]
zStar.lower = quantile(ZBoot, 0.025)
zStar.upper = quantile(ZBoot, 0.975)
round(mean(samSpotifyEnergy) - c(zStar.upper, zStar.lower) * se.avg(samSpotifyEnergy), 2)
```

