---
title: "Question 3"
output:
  pdf_document: default
  html_document: default
---

##### What is a Sample?
Sometimes, it's not possible to calculate an attribute for the entire population because of various reasons - such as the entire population is too large, or it's not completely available to us, or even if the attribute is too complex.
In the such scenarios, we would take a subset $S$ of $n << N$ units of the population to calculate the attribute. The result that we get based on the sample is an **estimate** $a(S)$ of the population attribute $a(P)$ such that, $$a(S)=a(\hat{P})$$  While using a sample, we consider 2 important aspects - **Sample Error and Fisher consistency** 


##### Sample Error
We can define sample error as the difference between the actual value of the estimate $a(S)$ and the quantity being estimated (or the true value $a(P)$).
$$ Sample Error = a(S) - a(P) $$

Let's look at example. We will look at the differences between some attributes. The sample we will use here would be of size $N$ = 40.

```{r}
temp <- read.csv("temperature.csv", header = TRUE)
set.seed(341)
temp.jan <- temp$JAN
s = sample(length(temp$JAN), 40)
c(mean(temp.jan[s]) - mean(temp.jan), median(temp.jan[s]) - median(temp.jan),
sd(temp.jan[s]) - sd(temp.jan), IQR(temp.jan[s]) - IQR(temp.jan))

```

We prefer an attribute with low sampling errors. 

##### Fisher Consistency

Ronald A. Fisher identified a consistency that, when the sample reaches the size of the population, the sample error approaches zero/non-existent. This consistency is known as Fisher consistency. 

Now, we will look at the same differences between the attributes as shown above but with different sample
sizes 

```{r}
s <-sample(length(temp.jan),60)
c(mean(temp.jan[s]) - mean(temp.jan), median(temp.jan[s]) - median(temp.jan),
sd(temp.jan[s]) - sd(temp.jan), IQR(temp.jan[s]) - IQR(temp.jan))
s <-sample(length(temp.jan),166)
c(mean(temp.jan[s]) - mean(temp.jan), median(temp.jan[s]) - median(temp.jan),
sd(temp.jan[s]) - sd(temp.jan), IQR(temp.jan[s]) - IQR(temp.jan))

```

We observe from the above results that as the sample size $N$ increases, the difference between the sample
estimate and the population decreases. Hence as Sample Size -> Population, the sample error -> 0

##### All possible samples
In a population $P$ of size $N$ and a sample of size $n$, the number of different possible samples $S$ are ${N \choose n}$

Hence, we can calculate the attribute for each possible sample. Let's try it out in the following example. We will find the average length of sharks of all possible samples.

```{r}
sharks <- read.csv("sharks.csv")
popSharks <- rownames(sharks)
sharks = na.omit(sharks)
popSharksAus <- popSharks[sharks$Australia == 1]
samples <- combn(popSharksAus, 5)
m <- ncol(samples)
avg.Samples <- apply(samples, MARGIN=2, FUN=function(s) {
mean(sharks[s, "Length"])
})
# Printing the average in the first 10 samples, and the last sample
avg.Samples[c(1:10, m)]

```

Let us also look at the sample error for all possible samples

```{r,echo=FALSE}
sampleErrors <- avg.Samples - mean(sharks[popSharksAus,"Length"])
sampleErrors[c(1:5, m)]
```

Now, let's look at it graphically. 
```{r}
par(mfcol=c(1,2))
av <- mean(sharks[popSharksAus,"Length"])
# Histogram of sample mean length
hist(avg.Samples, xlab="Average shark length", main="All possible sample averages", col="grey")
abline(v=av, col="blue", lwd=3, lty=2)
# Histogram of sample errors
hist(sampleErrors, xlab="Sample error", main="All possible sample errors", col="grey")
abline(v=0, col="red", lwd=3,lty=2)
```

##### Consistency & Effect of Sample Size
As the sample size increases, the sample approaches the population and attribute values will concentrate around the true value. 
We can quanity this concentration by, 
$$ | a(S) - a(P)| = \left|\frac{1}{n} \sum_{u \in S}y_u - \frac{1}{N}\sum_{u \in P}y_u\right| < C, \  for \ c > 0 $$
The above stands for the absolute difference between the sample attribute and the population attribute. 

Now, for each n, the possible set of samples will be 
$$ P_S(n) = \{S : S \subset P \ and \  |S| =n\}
$$

For any $c>0,$
$$ P_a(c,n) = \{S : S \subset P_S(n) \ and \ |a(S) - a(P)| < c\}
$$
and we define the proportion as  $$P_a(c,n) = \frac{|P_a(c,n)|}{|P_S(n)|}$$