---
title: "Question 1"
author: "Jin Barai"
output:
  pdf_document: default
  html_document: default
---

```{r}
library(carData)
data(TitanicSurvival)
Titanic = na.omit(TitanicSurvival)
Titanic = Titanic[Titanic$sex == "female", ]
Titanic$survived1 = as.numeric(Titanic$survived == "yes")
head(Titanic)
```

```{r}
TitanicSample <- c(57, 349, 147, 21, 310, 293, 62, 96, 304, 31, 258, 5,
381, 295, 180, 354, 346, 245, 88, 1, 383, 188, 333, 216, 351)
```


```{r}
popSize <- function(pop) {nrow(as.data.frame(pop))}
sampSize <- function(samp) {popSize(samp)}

createInclusionProbFn <- function(pop, sampSize) {
  N <- popSize(pop)
  n <- sampSize
  function(u) { n/N }
}

createJointInclusionProbFn <- function(pop, sampSize) {
  N <- popSize(pop)
  n <- sampSize
  function(u,v) { 
    ## Note that the answer depends on whether u and v
    ## are the same or different
    if (u == v) {n/N} else {(n * (n-1)) / (N * (N-1))}
  }
}

createHTestimator <- function(pi_u_fn) {
  function(samp, variateFn) {
    Reduce(`+`, 
           Map(function(u) {variateFn(u)/ pi_u_fn(u)}, samp),
           init = 0
    )
  }
}

createHTVarianceEstimator <- function(pop, pi_u_fn, pi_uv_fn) {
  function(samp, variateFn) {
    Reduce(`+`,
           Map(function(u) {
             pi_u <- pi_u_fn(u)
             y_u <- variateFn(u)
             Reduce(`+`, 
                    Map(function(v) {
                      pi_v <- pi_u_fn(v)
                      pi_uv <- pi_uv_fn(u, v)
                      y_v <- variateFn(v)
                      Delta_uv <- pi_uv - pi_u * pi_v
                      result <- (Delta_uv  * y_u * y_v) 
                      result <- result/(pi_uv * pi_u * pi_v)
                      result
                    }, 
                    samp),
                    init = 0) 
           },
           samp
           ),
           init = 0)
  }
}
```

\

##### (a)

```{r}
n = 25 
N = nrow(Titanic)
inclusionProb <- createInclusionProbFn(1:N, sampSize = n)
inclusionJointProb <- createJointInclusionProbFn(1:N, sampSize = n)

titanicHTestimator <- createHTestimator(inclusionProb)
HTVarianceEstimator <- createHTVarianceEstimator(1:N, 
                         pi_u_fn = inclusionProb, 
                        pi_uv_fn = inclusionJointProb)
```

```{r}
createVariateFnAvgAge <- function(popData, variate1, N=1, y=NULL) {
  function (u) { popData[u, variate1]/N }
}
TitanicAvgAge <- createVariateFnAvgAge(Titanic, "age", N=N)
```

The HT estimate of the average age is
```{r}
titanicHTestimator(TitanicSample, TitanicAvgAge)
```

An estimate of the variance or the standard error is (square root of the variance estimate)
```{r}
sqrt(HTVarianceEstimator(TitanicSample, TitanicAvgAge))
```

##### (b)

```{r}
createvariateFnNy <- function(popData, variate1, N=1, y=NULL) {
  function (u) { (popData[u, variate1] <= y )/N}
}

propTitanicAge25 =createvariateFnNy(Titanic, "age", N=N, y=25)
```


```{r}
pair <- round( c( titanicHTestimator(TitanicSample, propTitanicAge25), 
   sqrt(HTVarianceEstimator(TitanicSample, propTitanicAge25)) ), 3)
```

The HT estimate of the proportion of age less than or equal to 25 is
```{r}
pair[1]
```
The standard error of the above estimate 
```{r}
pair[2]
```

##### (c)

The HT estimate of the proportion of age less than or equal to 45 is on the left and the standard error is on the right: 

```{r}
propTitanicAge45 =createvariateFnNy(Titanic, "age", N=N, y=45)

round( c( titanicHTestimator(TitanicSample, propTitanicAge45), 
   sqrt(HTVarianceEstimator(TitanicSample, propTitanicAge45)) ), 3)
```

##### (d)

```{r}
yseq = c(0, sort(Titanic$age[TitanicSample]), 80)

cdf.estimate.sd = sapply(yseq, function(y) {  
  propAge <- createvariateFnNy(Titanic, "age", N=N, y=y)
  
  c( titanicHTestimator(TitanicSample, propAge), 
   sqrt( round(HTVarianceEstimator(TitanicSample, propAge), 14) ) )
   } )
  

par(mfrow=c(1,2) )
plot(yseq, cdf.estimate.sd[1,], type='s', ylab="Proportion",
     xlab="Age", 
     main="Estimate of the quantile or cdf")
plot(yseq, cdf.estimate.sd[2,], type='s', ylab="Standard Error",
     xlab="Age", 
     main="Estimate of the Variance of the cdf")
```

##### (e)

```{r}
plot(yseq, cdf.estimate.sd[1,], type='s', ylim=c(0,1),
     ylab="Proportion", xlab="Age", 
     main="Estimate of the quantile or cdf")

cdf.lower = cdf.estimate.sd[1,] - 2*cdf.estimate.sd[2,]
cdf.upper = cdf.estimate.sd[1,] + 2*cdf.estimate.sd[2,]

lines(yseq, cdf.lower, type='s',col=2, lty=2)
lines(yseq, cdf.upper, type='s',col=2, lty=2)

```

##### (f) 

The HT estimate of the proportion of females who survived is on the left and standard error is on the right:

```{r}
createvariateFnNyBool <- function(popData, variate1, N=1, y=NULL) {
  function (u) { (popData[u, variate1] == y )/N}
}

propTitanicSurvived <- createvariateFnNyBool(Titanic, variate1="survived1", y=1, N=N)
round( c( titanicHTestimator(TitanicSample, propTitanicSurvived), 
   sqrt(HTVarianceEstimator(TitanicSample, propTitanicSurvived)) ), 3)

```

##### (g) 

The HT estimate of the proportion of females in 1st class is on the left and standard error is on the right:

```{r}
propTitanicFirstClass<- createvariateFnNyBool(Titanic, variate1 = "passengerClass",y="1st", N=N)
round( c( titanicHTestimator(TitanicSample, propTitanicFirstClass), 
   sqrt(HTVarianceEstimator(TitanicSample, propTitanicFirstClass)) ), 3)

```


