---
title: "Question 4"
author: "Jin Barai"
output:
  pdf_document: default
  html_document: default
---

##### Introduction

Suppose we have the population $\mathcal{P}=\{\mathcal{P}_{1},\mathcal{P}_{2}\}$. We can now compare some attributes across the sub-populations of the above population. We need to understand if the difference between attributes of two sub-populations - $a(\mathcal{P}_{1})$ and $a(\mathcal{P}_{2})$  is unusual or expected. 

We use a test of significance to understand this, relative to *randomly mixed* sub-populations. 

##### Test of Significance Algorithm 

1. Stating the **null hypothesis** denoted by $H_{0}$: $\mathcal{P}_{1}$ and $\mathcal{P}_{2}$ are drawn from the same population
2. Construct a discrepency measure $D = D(\mathcal{P}_{1}, \mathcal{P}_{2})$ where the large values indicate evidence **against** $H_{0}$. The discrepency measure is also referred to as **test statistic**.
3. Calculate the **observed discrepancy** $d_{obs} =  D(\mathcal{P}_{1}, \mathcal{P}_{2})$ 
4. Shuffle the sub-populations $M$ times and calculate the **observed p-value** given by:
     $$ p-value = Pr(D \ge d_{obs} \mid H_{0}\ is\ true) \approx  \frac 1 M\sum _{i=1}^{M}I(D(\mathcal{P}^{*}_{1,i},\mathcal{P}^{*}_{2,i}) \ge d_{obs}) $$
which basically means finding the probability that a randomly shuffled sub-population has a discrepency measure ($D$) **at least** as large as what we observed ($d_{obs}$), given that $H_{0}$ is true. The smaller the $p$-value, the greater the evidence against the null hypothesis. 


*Note*: In order to calculate the $p$-value *exactly*, one must consider all $N_{1}+N_{2} \choose N_{1}$ = $N_{1}+N_{2} \choose N_{2}$ possible permutations, but since it is too many permutations to consider we just use M of them.
     
##### Example

Let's use the TitanicSurvival data to find an estimate of the significance levels using test of significance algorithm. 
The two sub-populations would be the $\mathcal{P}_{male}$ = survived male passengers and $\mathcal{P}_{female}$ = survived female passengers. The attribute would be the average age.

$H_{0}$: The two populations ($\mathcal{P}_{male}$,$\mathcal{P}_{female}$) are randomly drawn. 

```{r, echo=FALSE}
mixRandomly <- function(pop) {
  pop1 <- pop$pop1
  n_pop1 <- nrow(pop1)
  pop2 <- pop$pop2
  n_pop2 <- nrow(pop2)
  mix <- rbind(pop1, pop2)
  select4pop1 <- sample(1:(n_pop1 + n_pop2), n_pop1, replace = FALSE)
  new_pop1 <- mix[select4pop1, ]
  new_pop2 <- mix[-select4pop1, ]
  list(pop1 = new_pop1, pop2 = new_pop2)
}
```

```{r, echo=FALSE}
getAveDiffsFn <- function(variate) {
  function(pop) {
    mean(pop$pop1[, variate]) - mean(pop$pop2[, variate])
  }
}
getSDRatioFn <- function(variate) {
  function(pop) {
    sd(pop$pop1[, variate])/sd(pop$pop2[, variate])
  }
}
```

```{r, echo=FALSE}
library(carData)
data(TitanicSurvival)
Titanic = na.omit(TitanicSurvival)
Titanic$survived1 = as.numeric(Titanic$survived == "yes")
```

```{r}
Titanic = Titanic[Titanic$survived == "yes",]
pop <- list(pop1 = subset(Titanic, Titanic$sex == "male"), pop2 = subset(Titanic, Titanic$sex=="female"))
```

```{r}
# The difference in average age of the subpopulations
diffAveAge <- getAveDiffsFn("age")
# The ratio of the standard deviations between the subpopulations 
ratioSDAge <- getSDRatioFn("age")
round(c(diffAveAge(pop), ratioSDAge(pop)),3)
```

Randomly mixing the two sub-populations and calculating the above summaries. We use a sample of $M$=5000 shuffled pairs in this case. 
 
```{r}
set.seed(341)
mixedPop <- mixRandomly(pop)
diffAge <- sapply(1:5000, FUN = function(...) {
  diffAveAge(mixRandomly(pop))
})
round(c(diffAveAge(mixedPop), ratioSDAge(mixedPop)), 3)
```

Finding the $p$-value
```{r}
sum(abs(diffAge) >= abs(diffAveAge(pop))) / length(diffAge)
```
```{r, echo=FALSE}
plot.margin = c(1,1,1,1)
par(mfrow = c(1,2))
set.seed(341)
agepair <- sapply(1:5000, FUN = function(...) {
  tmixpop = mixRandomly(pop)
  c(diffAveAge(tmixpop), ratioSDAge(tmixpop))
})
hist(agepair[1,], breaks = "FD", main = "Randomly Mixed Populations", xlab = "difference in averages of ages", col = "lightgrey")
abline(v = diffAveAge(pop), col = "red", lwd = 2)
hist(agepair[2,], breaks = "FD", main = "Randomly Mixed Populations", xlab = "ratio of standard deviations",col = "lightgrey")
abline(v = ratioSDAge(pop), col = "red", lwd = 2)
```

Conclusions:
  
* $p-value \approx  0.0704$ for the null hypothesis $H_{0}$

* From the $p$-value we can conclude that there is **weak evidence** against $H_{0}$

* Hence, we have very weak evidence against the null hypothesis that the pair $(\mathcal{P}_{male}$ ,$\mathcal{P}_{female})$ was randomly drawn  

##### Remarks

* The observed $p$-value doesn't measure evidence **in favour** of $H_{0}$. 
* The observed $p$-value provides a common (probabilistic) scale on which to measure the **evidence against null hypothesis**
* A test of significance **neither accepts nor rejects a null hypothesis**, it provides a measure of evidence against it. 
